Peer-to-Peer Permissionless Consensus via Reputation

### 2023-08-11 freechains/rebuttal.md

# Rebuttal Letter

We would like to thank the reviewers for the thoughtful comments, questions,
and criticism.

We accommodate all clarifications and answers in the final version of the
paper.

The link that follows shows the *diff* to reach the final version:

- TODO

Each comment is addressed individually as follows.

@@

## Reviewer 1

> Quanto a solução, senti falta de alguma discussão sobre as limitações.
> Em especial, a falta de finalidade, i.e., um post pode ser excluído devido
> aos forks (no momento do merge).
> Este post poderia ser de um usuário legitimo ? quanto isso ocorre e quais as
> implicações?

Sim, para isso, ele teria que escrever em um fork "obscuro", desconectado do
resto dos usuários legítimos.
Dessa forma, ao ser reconectado, esse fork poderia sim ter posts removidos.
A única implicação é que ele teria que postar novamente a mesma mensagem.

@@

No meio da Seção 2.1, essa possibilidade é levantada, mas não com tanta ênfase:

> While applying the branches in order, if any post operation fails, all
> remaining posts are rejected and removed from the DAG, as if they never
> existed.

Em termos práticos, os usuários legítimos se conectam mais diretamente ao
longo do tempo, o que deveria ser suficiente para evitar essa situação.
Para a versão final, nós adicionamos o seguinte trecho:

> Legitimate posts in secondary branches might be removed due to merges, which
> requires that users repost the messages. Nevertheless, we expect that
> legitimate users gossip among themselves more often, thus preventing this
> situation.

@@

Sim, se o usuário malicioso dispor de 1 rep, ele pode escrever propositalmente
em vários forks.
Nesse caso, a reputação prévia dele vai contar igualmente em todos os merges,
se anulando e, portanto, não influenciando no resultado final do merge.

Em caso de empate, que pode acontecer mesmo sem usuários maliciosos, o hash do
fork será critério considerado por todos igualmente.
Esse critério é descrito no final da Seção 2.2:

> ...otherwise, branches use the arbitrary criteria of lexicographical order of
> the post hashes immediately after the common prefix.

Como um possível ataque, o usuário malicioso poderia escrever em vários
branches antes de usuários legítimos, de forma que os merges acabariam por
remover mensagens legítimas.
Essa situação, que complementa o comentário anterior, obrigaria os usuários a
repostarem as mensagens.
Da mesma forma, nós entendemos que os usuários legítimos estarão mais
conectados, inviabilizando esse tipo de ataque.

@@

> It is strange that known techniques like EigenTrust are not even mentioned,
> even though we are in the blockchain domain.

TODO

> We do not know if the results are good or bad because there was no comparison
> to the state of the art or baseline.

The main goal of our experiments in Section 3.1 is to stress the consensus
algorithm to show that the mechanism is feasible, i.e., that it has a
reasonable performance to be used in practice.
The good result is that the protocol works in a permissionless context,
regardless of the slower performance.

@@

We included the following remarks:

> We consider Sybils to be groups of throw-away identities and machines with
> no previous reputation in the forums.
> ...
> In this sense, having one or hundreds of machines in the network does not
> affect the capacity to write into the forums.
> ...
> In our analogy with Bitcoin, Sybil attacks would require to add human
> resources, instead of CPU power.

# Reviewer 3

> It's important to note that the reputation system does not enforce "good"
> human behavior but provides a transparent and quantitative way for users to
> understand forum evolution and act accordingly. 
> ...
> I am also concerned that the content removal is too weak to work in practice.
> For example, a Forum may be taken over by extremists, and they can upvote
> illegal content. That would require removal due to legal process, or face
> blocking at network level.

This is all true, and is the reason why we close the paper with your remark
above.
It is not even required to be "taken by extremists", but only "created by and
for extremists".
But we also believe that a protocol that is tamper proof, and provides
persistence and data transparency can actually support law enforcement.

@@

> Maybe a way to go is to create a pool of super-reputations that could answer
> to injunctions and administratively remove the content and a measure to avoid
> illegal content being broadcasted in subverted grupus.

For extreme conflicts, in which content removal is too weak, the protocol also
provides hard forks, as discussed in the middle of Section 2.2:

> More than simple numeric disputes, a hard fork represents a social conflict
> in which reconciling branches is no longer possible.

> The paper is too descriptive, so reading sometime is tiresome.
> ...
> Reading could be improves by using some figures to explain ideas. Sometimes
> it is difficult to follow the text.

We opted for a full description of the protocol, but we recognize that the
paper became a bit dense.
It is hard to modify the whole text without more specific suggestions, but we
removed some details of Freechains from Section 3.

### 2023-08-13 freechains/freechains.tex
...

There are some other relevant considerations about forks and merges:
%
Peers that first receive branches with less reputation need to reorder all
posts after the forking point.
This might involve removing content in the end-user software.
This behavior is similar to Bitcoin's blockchain reorganization, when peers
detect new longest chains. % and disconsiders old blocks.
%
Likewise, peers that first saw branches with more reputation just need to
append the other branch, with no reordering at all.
%and do not need to recompute anything.
This behavior is expected to happen in the majority of the connected network.
%
Legitimate posts in secondary branches might be removed due to merges, which
requires that users repost the messages.
Nevertheless, we expect that legitimate users gossip among themselves more
often, preventing this situation.
%
Note that unlike Bitcoin, forks are not only permitted but encouraged due to
the local-first software principle~\cite{p2p.local}, in which networked
applications can work locally while offline.
However, the longer a peer remains disconnected, the more conflicting
operations it may see, and the higher are the chances of posts reordering when
rejoining.

@@

Many other systems have been proposed for decentralized content
sharing~\cite{p2p.survey,p2p.ecosystem}.
Here we consider the classes of publish-subscribe, federated, and P2P
protocols.

@@

>>>>

There are some other relevant considerations about forks and merges:
%
Peers that first receive branches with less reputation need to reorder all
posts after the forking point.
This might involve removing content in the end-user software.
This behavior is similar to Bitcoin's blockchain reorganization, when peers
detect new longest chains. % and disconsiders old blocks.
%
Likewise, peers that first saw branches with more reputation just need to
append the other branch, with no reordering at all.
%and do not need to recompute anything.
This behavior is expected to happen in the majority of the connected network.
%
Legitimate posts in secondary branches might be removed due to merges, which
requires that users repost the messages.
Nevertheless, we expect that legitimate users gossip among themselves more
often, thus preventing this situation.
%
Note that unlike Bitcoin, forks are not only permitted but encouraged due to
the local-first software principle~\cite{p2p.local}, in which networked
applications can work locally while offline.
However, the longer a peer remains disconnected, the more conflicting
operations it may see, and the higher are the chances of posts reordering when
rejoining.

@@

%Many other systems have been proposed for decentralized content
%sharing~\cite{p2p.survey,p2p.ecosystem}.
%Here we consider the classes of publish-subscribe, federated, and P2P
%protocols.

@@

Some works for P2P file sharing propose to account the reputation of peers to
form a web of trust based on their behavior
history~\cite{p2p.rep.wang,p2p.rep.eigentrust}.
However, there are three key aspects in terms of scope that distinguishes our
reputation system: its focus on the contents, its subjective nature, and its
dependency on consensus.
First, it is the actual contents that are stored and evaluated in the forums,
with the identities being a secondary attribute.
Second, content evaluation is subjective, as well as any decision to revoke
posts or to fork forums.
Third, at any given point in time, peers must agree on a common forum DAG
prefix with global and deterministic accuracy to act identically as a whole.
In this case, consensus is fundamental, since a subtle off-by-one discrepancy
in reputation creates an irreconcilable fork in the network.
Achieving consensus on a permissionless network is the key contribution of this
work.
